\documentclass[10pt]{article}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}

\usepackage{cmbright}
\usepackage[T1]{fontenc}

\usepackage{multicol}
\setlength\columnsep{25pt} 

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{tikz}
\usepackage{graphicx}
\graphicspath{  {./images/} }
\setlength{\parindent}{0pt}
\usepackage{changepage}
\usepackage{physics}
\usepackage{derivative}
\usepackage{bm}

\addtolength{\topmargin}{-.5in}	
\addtolength{\textheight}{0.75in}
\addtolength{\oddsidemargin}{-.75in}
\addtolength{\textwidth}{1.5in}

\makeatletter
\newcommand*\bigcdot{\mathpalette\bigcdot@{.5}}
\newcommand*\bigcdot@[2]{\mathbin{\vcenter{\hbox{\scalebox{#2}{$\m@th#1\bullet$}}}}}
\makeatother

\DeclareMathOperator{\di}{d\!}
\newcommand*\Eval[3]{\left.#1\right\rvert_{#2}^{#3}}

\newcommand{\uvec}[1]{\boldsymbol{\hat{\textbf{#1}}}}
\newcommand{\vr}[1]{\textbf{#1}}

\newcommand{\thus}[0]{\; \; \longrightarrow \; \;}

\newcommand{\lag}{\mathcal{L}}
\newcommand{\ham}{\mathcal{H}}

\title{Gravitational Wave Research Overview}
\author{Ryan Liu}
\date{Last updated: \today}

\begin{document}

\maketitle

\begin{multicols*}{2}


\section{Research Proposal}

\subsection{Research Question}

\textbf{Question:} How far out in distance can we detect the gravitational wave emission of a black hole merger given a specific signal-to-noise ratio (SNR)? \\

\textbf{Initial Assumptions:}
\begin{itemize}
    \setlength\itemsep{0pt}
    \item The two black holes are of equal mass.
    \item The black holes are not spinning.
    \item The orientation of the detectors relative to the GW signal is random. 
\end{itemize}

\subsection{Initial Objectives}

\begin{enumerate}
    \item Determine the redshift effect on gravitational waves on a cosmological scale. Create a function that generates waveforms based on mass and distance, factoring in redshift. 
    \item Create a function that generates random noise indistinguishable from O3 aLIGO data. 
    \item Create a function that inserts waveforms into aLIGO data, then determines the average SNR that is generated by the signal. 
    \item Create a function that determines the distance threshold for a particular SNR value and particular mass. Create a visualization for the results of this function. 
    \item Factor in any additional effects. 
\end{enumerate}

\section{Summary: April}

\subsection{Cosmological Redshift}

A function was created to generate the waveform of a black hole merger given the masses of the two black holes and the distance of observation. The redshift effect was added based on Hubble's law. 

\subsection{Random Noise Generator}

To simulate the detection of a black hole merger by LIGO, a function was created to overlay the redshifted signal with noise based on various template PSDs. Then, the signal will be "found" with a particular waveform template at a particular SNR. 

\subsection{SNR Calculation}

To answer the research question, the functions from the previous two sections were used to look at the impact of distance of observation, imprecision of the template waveform, LIGO sensitivity, black hole mass, and random noise on the SNR of detection. 

\section{Summary: May}

\subsection{Recalculated Redshift}

Because a linear approximation of redshift is only accurate at small distances, a more refined redshift approximation as a function of $\Omega_m$ and $\Omega_v$ was calculated. These values were checked with the AstroPy package, using the Planck 2018 values for cosmological constants. 

\subsection{SNR Expectation Value Function}

In order to remove the effect of noise on determining the SNR of a BH signal, the function created in Section (2.2) was modified to remove noise. Additionally, to speed up computation the calculations were moved into the frequency spectrum. 

\subsection{SNR Expectation Simulations}

Similar to Section (2.3), the SNR was calculated for black hole mergers of varying masses, distances, sensitivity curves, and templates. The maximum distance that a black hole merger can be observed as a function of frequency was also determined. 

\subsection{SNR Expectatation vs. Noise}

The results of SNR calculations using the functions from Section (2.2) and (3.2) were compared to determine whether they agree with one another. Surprisingly, the expectation value appears to be consistently higher than the median SNR when noise is added for the same BH merger signal. 

\subsection{Distance Estimation Function}

Because the functions created in (2.2) and (3.2) determine SNR as a function of multiple parameters (including distance) and cannot be inverted, in order to answer the original research question a function was created using the concept of parameter estimation to determine distance as a function of the desired SNR threshold. 

\subsection{Distance Estimation Simulations}

The maximum distance a black hole merger can be detected as a function of mass, LIGO sensitivity, and threshold SNR was calculated in this section. Due to inaccuracies in function convergence, a more detailed look into to results of (3.3) was also conducted. 

\subsection{Spin Simulations}

The effect of spin on GW signals and their relative SNR were investigated. 

\section{Summary: June-July}

\subsection{SNR Expectation Value Function}

The entire expectation value function was moved into the frequency domain, including generation of the waveform. Additionally, the function was amended to scale with luminosity distance rather than comoving distance. All sources of noise due to conversion error were erased, and the function runs around 40 times faster. 

\subsection{Distance Estimation Simulations}

With the speed improvement over (3.5), the distance estimation function was run for all three design sensitivities at an SNR of 6, 8, and 10. For an SNR of 8, the maximum distance of observation as a function of total mass and mass ratio was calculated. 

\subsection{Merger Rate Calculations}

The BBH merger rate and total mergers as a function of redshift were calculated for non-evolving and modified power law distributions. 

\subsection{Spin Distribution Calculations}

Using IMRPhenomPv2, the effects of spin in all three dimensions ($x$, $y$, and $z$) were investigated. Distance estimation simulations were run for varying component spins, effective spins, and precession spins, and various Gaussian spin distributions were calculated. 

\subsection{Mass Distribution Calculations}

The mass distribution of binary BBH mergers (primary mass and mass ratio) were determined using several modified power law distributions between $\approx 3$ to $90 M_\odot$. 

\subsection{Sky Location Calculations}

The effects of sky location (celestial coordinates), inclination, and detector orientation on the expectation SNR of BBHs were investigated. Additionally, the antenna pattern volume as a function of redshift was calculated.

\subsection{Detection Rate Calculations}



\section{Final Paper Outline}

\textbf{Title:} Determining Binary Black Hole Detection Rates for Second- and Third-Generation Gravitational-Wave Observatories

\begin{enumerate}
    \setlength{\itemsep}{0pt}
    \item Introduction
    \begin{enumerate}
        \item Brief overview of LIGO detections and GW pouplations 
        \item Overview of contents of paper, outlining goals, parameters, and results
    \end{enumerate}
    \item Estimation of merger rate density
    \begin{enumerate}
        \item Mass-redshift-spin model
        \item Subsections for mass, redshift, spin, and location distributions
    \end{enumerate}
    \item Calculation of LIGO detection ranges
    \begin{enumerate}
        \item SNR calculations
        \item Parameter estimation model
        \item Results on number of mergers detected
        \item Population distribution of mergers
    \end{enumerate}
    \item Conclusions
\end{enumerate}

\section{Questions}

\begin{itemize}
    \item Is there a preferred model for the mass distribution / spin distribution / merger rate that should be used or considered? 
    \item How valid are extrapolations of the above models to the high redshifts that can be detected by Cosmic Explorer?
    \item How can we iterate through all of the $(m_1, q, \chi_\text{eff}, \chi_\text{p})$ bins efficiently? Should I (and how) determine an analytic function for the SNR? 
    \item What will eventual publication of the paper entail? Rough schedule? 
\end{itemize}






\end{multicols*}
\end{document}
